---
title: "Self-Contact in the Crypt"
author: "Axel Almet"
date: "23 May 2018"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(cowplot)
library(R.matlab)
library(latex2exp)
library(RColorBrewer)
library(ggpubr)
library(plotly)
```

# Current model
The general form of the kinematic and mechanics of the rod read as:

\begin{align}
&\mathbf{r}' = \alpha\gamma\mathbf{d}_3,\\
&\mathbf{n}' + \gamma\mathbf{f} = \mathbf{0},\\
&\mathbf{m}' + \mathbf{r}'\times\mathbf{n} = \mathbf{0}.
\end{align}
The general form of the foundation can be written as
$$ \mathbf{f} = k\frac{\sigma}{\Delta}(\pmb{\rho} - \mathbf{r}), \qquad\mbox{ where }\qquad \Delta = \| \pmb{\rho} - \mathbf{r}\|.$$
Writing $\rho = S_0\mathbf{e}_x$ and projecting these equations into the Cartesian plane yields
\begin{align}
&x' = \alpha\gamma\cos\theta, \qquad y' = \alpha\gamma\sin\theta,\\
&F' = k\frac{\sigma}{\Delta}(x - S_0), \qquad G' = k\frac{\sigma}{\Delta}y,\\
&m' = \alpha\gamma(F\sin\theta - G\cos\theta).
\end{align}

These equations are supplemented by two constituve relations for bending and stretching, respectively:

$$
m = E_b\gamma^{-1}\theta', \qquad \alpha = 1 + E_s^{-1}(F\cos\theta + G\sin\theta).
$$

The rod is assumed to be clamped at both ends:

$$
x(0) = 0, \quad x(L_0) = L_0, \qquad y(0) = y(L_0) = 0, \qquad \theta(0) = \theta(L_0) = 0.
$$

## Contact at a point: jump force
Suppose that contact within the rod occurs between two points, $(x(s_1), y(s_1))$ and $(x(s_2), y(s_2))$. If we assume the contact is symmetric about $x = L_0/2$, then we can write
$$ s_1 = \frac{L_0}{2} - s_c, \qquad s_2 = \frac{L_0}{2} + s_c.$$
At the point of contact, we modify the external force as such:
$$ \mathbf{f} = k\frac{\sigma}{\Delta}(\pmb{\rho} - \mathbf{r}) -\sigma_c\delta(s - s_1)\mathbf{e}_x + \sigma_c\delta(s - s_2)\mathbf{e}_x.$$
Projecting this into $x-y$ coordinates yields a modified horizontal force balance equation:
$$ F' = k\frac{\sigma}{\Delta}(x - S_0) + \sigma_c(\delta(s - s_1) - \delta(s - s_2)).$$
As we are working with Dirac delta functions, we must ensure continuity of the horizontal force. Integrating the force balance about $s = s_1$ and $s = s_2$ allows us to obtain the necessary jump conditions:
\begin{align}
\left[ F\right]^{s = s_1^+}_{s = s_1^-} &= \int^{s_1^+}_{ s_1^-}k\frac{\sigma}{\Delta}(x - S_0)ds + \sigma_c,\\
\left[ F\right]^{s = s_2^+}_{s = s_2^-} &= \int^{s_2^+}_{ s_2^-}k\frac{\sigma}{\Delta}(x - S_0)ds - \sigma_c.
\end{align}

With our symmetry assumption, both $s_1$ and $s_2$ are determined by $s_c$. Therefore, the challenge is to solve for $s_c$ and $\sigma_c$. Considering the geometry of the rod at $s = s_c$ gives two extra conditions to account for these unknown variables:
$$ x(s_1) = \frac{L_0}{2} - w, \qquad \theta(s_1) = \frac{\pi}{2}. $$
Here, $w \ge 0$, is a fixed parameter chosen to represent the width of contact. The computational difficulty is clearly incorporating these constraints into the current framework.

## Contact at a point: normal pressure (trapped fluid)

Another way to envision self-contact is to imagine that after the contact, some type of fluid (e.g. mucus) becomes trapped within the loop that is created. This in turn exerts a normal pressure on the crypt such that this area is maintained.

By Green's theorem, the area, $A_0$, between the contact points $s_1$ and $s_2$, is given by 

$$ A_0 = \int^{s_2}_{s_1} x(s)y'(s)ds. $$

This constant area constraint can be accounted for via a Lagrange multiplier in the energy functional. Minimising this new functional via the Calculus of Variations gives rise to new force balance equations:

\begin{align}
F' &= k\frac{\sigma}{\Delta}(x - S_0) + \sigma_cy', 
\\G' &= k\frac{\sigma}{\Delta}y - \sigma_cx'.
\end{align}

Hence, the contact force is given by the vector

$$ \mathbf{f}_\mathrm{new} = \sigma_c(y'\mathbf{e}_x - x'\mathbf{e}_y). $$
As this contact force is parallel to the normal director basis vector $\mathbf{d}_1 = \sin\theta\mathbf{e}_x - \cos\theta\mathbf{e}_y$, we see that this area constraint exerts a normal pressure on the force balance. The force magnitude $\sigma_c$ is unknown and is to be determined in the resulting numerical solution.

## Reaction force  over a region

Suppose now that we have contact over an interval. On the left, contact occurs over $[s_1, s_2]$, and $[s_3, s_4]$ on the right. Under symmetry assumptions similar to the point case, one can assume

$$ s_1 = \frac{L_0}{2} - s_{c_1},  \quad s_2 = \frac{L_0}{2} - s_{c_2},  \quad s_3 = \frac{L_0}{2} + s_{c_2}, \quad s_4 = \frac{L_0}{2} + s_{c_1}.$$

The body force $\mathbf{f}$ must now account for contact over an interval, and is modified as thus:
$$ \mathbf{f} = \frac{\sigma}{\Delta}(\pmb{\rho} - \mathbf{r}) -\sigma_c(s)\mathbb{1}_{(s_1, s_2)}\mathbf{e}_x + \sigma_c(s)\mathbb{1}_{(s_3, s_4)}\mathbf{e}_x.$$
Here, $\mathbb{1}_{(a, b)}$ denotes the indicator function over the interval $(a, b)$. Projecting into $x$ and $y$ coordinates now yields
$$ F' = k\frac{\sigma}{\Delta}(x - S_0) + \sigma_c(s)\mathbb{1}_{(s_1, s_2)} - \sigma_c(s)\mathbb{1}_{(s_3, s_4)}.$$
The continuity condition is now different in that, because $\sigma_c(s)$ is a distributed force, we can only solve up to an 'averaged' value of $\sigma_c(s)$. Nevertheless, the conditions are now
\begin{align}
\left[ F\right]^{s = s_2}_{s = s_1} &= \int^{s_2}_{s_1}k\frac{\sigma}{\Delta}(x - S_0)ds + \int^{s_2}_{s_1}\sigma_c(s)ds,\\
\left[ F\right]^{s = s_4}_{s = s_3} &= \int^{s = s_4}_{s = s_3}k\frac{\sigma}{\Delta}(x - S_0)ds - \int^{s_4}_{s_3}\sigma_c(s)ds.
\end{align}
As contact is over an extended region, we must now ensure continuity of the vertical force, $G$:
\begin{align}
\left[ G\right]^{s = s_2}_{s = s_1}  \int^{s_2}_{s_1}k\frac{\sigma}{\Delta}yds,\qquad
\left[ G\right]^{s = s_4}_{s = s_3} = \int^{s_4}_{ s_3}k\frac{\sigma}{\Delta}yds.
\end{align}
We now have six unknown variables to solve for: $\{s_1, s_2, s_3, s_4,  \int^{s_2}_{s_1}\sigma_cds, \int^{s_4}_{s_3}\sigma_cds\}$.  Assuming symmetrical contact halves this to three unknown variables, $\{s_1, s_2,  \int^{s_2}_{s_1}\sigma_cds\}$. Geometric consideration of the rod during contact give us the three extra conditions to solve for these variables:

$$ x(s) = \frac{L_0}{2} - w, \quad \theta(s) = \frac{\pi}{2}, \quad \theta'(s) = 0, \qquad \forall s \in [s_1, s_2].$$
NB. It is currently unclear what the appropriate change to the final contact condition is when curvature remodelling is introduced.

## Numerical Implementation
We implement these contact models using the multipoint functionality of MATLAB's bvp4c. For bvp4c to be happy, it seems that one needs to enforce continuity of $x$, $y$, $\theta$, $G$, and $m$, as well.

<!-- ## Results -->

<!-- ### Self-contact with jump force -->
<!-- ```{r plot crypt maxwell self contact, echo=FALSE, fig.align="center", warning=FALSE} -->
<!-- directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/LinearViscoelasticFoundation/Maxwell/" -->
<!-- values = "jumpcontact_Eb_0p5_sigmaE_2w_simplified_nu_0p16_k_0p02_L0_0p125_sigma_2w_etaK_0p16_initforce" -->

<!-- # Load the solutions -->
<!-- solsList <- readMat(paste(directory, "sols_", values, ".mat", sep="")) -->
<!-- # Load the times -->
<!-- timesList <- readMat(paste(directory, "times_", values, ".mat", sep="")) -->
<!-- stressesList <- readMat(paste(directory, "maxwellstresses_", values, ".mat", sep="")) -->

<!-- numSols <- length(solsList[[1]]) -->
<!-- solData <- NULL # Create data frame for solutions -->
<!-- stressData <- NULL -->
<!-- curvData <- NULL -->

<!-- # Some parameters -->
<!-- solCount <- 1 -->

<!-- # # Plot solutions -->
<!-- for ( i in seq(1, numSols - 4, solCount) ) -->
<!-- { -->
<!--   # Extract the solution, growth and time -->
<!--   sol <- solsList[[1]][[i]][[1]] -->
<!--   currentTime = timesList[[1]][[i]] -->
<!--   stresses <- stressesList[[1]][[i]][[1]] -->

<!--   # Update the solutions -->
<!--   solDataFrame <- data.frame(S=sol[2,], x=sol[3,], y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,], sC=sol[9,], -->
<!-- fC=sol[10,]) -->
<!--   solDataFrame$time = toString(currentTime) -->
<!--   solData <- rbind(solData, solDataFrame) -->
<!-- } -->

<!-- solPlot <- ggplot(data=solData, aes(frame=time)) + -->
<!--   geom_path(aes((15/(2*sqrt(3)))*x, -(15/(2*sqrt(3)))*y + 2*6*125, frame=time), size=0.75) +  -->
<!--   ylab("y") + xlab("x") +  -->
<!--   theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!-- scale_x_continuous(expand = c(0, 0), limits=c(0,126)) +  -->
<!-- scale_y_continuous(expand = c(0, 0), limits=c(720, 790)) -->

<!-- solPlot <- solPlot + geom_path(aes(125 - (15/(2*sqrt(3)))*x, -(15/(2*sqrt(3)))*y + 2*6*125, frame=time), size=0.75) -->

<!-- animatedPlot <- ggplotly(solPlot, redraw=TRUE, frame=100, transition=0) -->

<!-- animatedPlot -->

<!-- ``` -->

### Self-contact with constant bud area

```{r plot crypt maxwell self contact normal pressure, echo=FALSE, fig.align="center", warning=FALSE}
directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/LinearViscoelasticFoundation/Maxwell/"
values = "normalcontact_Eb_init_0p5_sigmaE_2w_nu_simple_init_0p16_k_0p02_L0_0p125_sigma_init_2w_etaK_0p16"

# Load the solutions
solsList <- readMat(paste(directory, "sols_", values, ".mat", sep=""))
# Load the times
timesList <- readMat(paste(directory, "times_", values, ".mat", sep=""))
stressesList <- readMat(paste(directory, "maxwellstresses_", values, ".mat", sep=""))

numSols <- length(solsList[[1]])
solData <- NULL # Create data frame for solutions
stressData <- NULL
curvData <- NULL

# Some parameters
solCount <- 1

# # Plot solutions
for ( i in seq(1, numSols, solCount) )
{
  # Extract the solution, growth and time
  sol <- solsList[[1]][[i]][[1]]
  currentTime = timesList[[1]][[i]]
  stresses <- stressesList[[1]][[i]][[1]]

  # Update the solutions
  solDataFrame <- data.frame(S=sol[2,], x=sol[3,], y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,], A=sol[9,],
A0=sol[10], sC=sol[11,], fC=sol[12,])
  solDataFrame$time = toString(currentTime)
  solData <- rbind(solData, solDataFrame)
}

# solPlot <- ggplot(data=solData, aes(frame=time)) +
#   geom_path(aes((15/(2*sqrt(3)))*x, -(15/(2*sqrt(3)))*y + 2*6*125, frame=time), size=0.75) +
#   ylab("y") + xlab("x") +
#   theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
# scale_x_continuous(expand = c(0, 0), limits=c(0,126)) +
# scale_y_continuous(expand = c(0, 0), limits=c(720, 790))
#
# solPlot <- solPlot + geom_path(aes(125 - (15/(2*sqrt(3)))*x, -(15/(2*sqrt(3)))*y + 2*6*125, frame=time), size=0.75)

 solPlot <- ggplot(data=solData, aes(frame=time)) +
   geom_path(aes((15/(2*sqrt(3)))*x, -(15/(2*sqrt(3)))*y, frame=time), size=0.75) +
   ylab("y") + xlab("x") +
   theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
scale_x_continuous(expand = c(0, 0), limits=c(0,126))

solPlot <- solPlot + geom_path(aes(125 - (15/(2*sqrt(3)))*x, -(15/(2*sqrt(3)))*y, frame=time), size=0.75)

animatedPlot <- ggplotly(solPlot, redraw=TRUE, easing='back', frame=100, transition=0)

animatedPlot

```