---
title: "Investigating Growth Laws"
author: "Axel Almet"
date: "5 September 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(cowplot)
library(R.matlab)
library(latex2exp)
library(RColorBrewer)
library(ggpubr)
library(plotly)
library(tikzDevice)
```

# Set-up
We model the crypt as a planar, growing, inextensible elastic rod attached to an underlying foundation, whose attachments remodel over time. Tissue growth is encapsulated by the growth stretch $\gamma$, modelling a local increase of the initial arc length $S_0$; that is, 
\begin{align}
\gamma = \frac{\partial s}{\partial S_0}.
\end{align}
The equations describing geometric constraints and balance of linear and angular momentum are given by:
\begin{align}
&x' = \gamma\cos\theta,\\
&y' = \gamma\sin\theta,\\
&n_x' = \gamma k(x - p_x), \qquad\mbox{ where }\qquad \dot{p}_x = \eta(x - p_x),\\
&n_y' = \gamma k(y - p_y), \qquad\mbox{ where } \qquad\dot{p}_y = \eta(y - p_y),\\
&\theta' = E_b\gamma m,\\
&m' = \gamma(n_x\sin\theta - n_y\cos\theta).
\end{align}
For simplicity, we assume the rod is clamped horizontally at the boundaries:
\begin{align}
x(0) = 0, \qquad x(L_0) = 0, \qquad y(0) = y(L_0) =  y_0, \qquad \theta(0) = \theta(L_0) = 0.
\end{align}

The parameter $L_0$ is the initial rod length, giving the horizontal interval of deformation for the crypt.

# Contributions to growth
A generalised evolution equation for $\gamma$ takes the form
\begin{align}
\frac{\dot{\gamma}}{\gamma} = H(s, S_0, W, n_1, n_3, \sigma, \Delta, \cdots), 
\end{align}
where $H$ can depend on various quantities, such as position ($S_0$ or $s$), chemical signals ($W$), stresses ($n_1$, $n_3$, $\sigma$), strains ($\Delta$) etc. Here, we discuss possible contributions to include in our growth law. 

All simulations here have been plotted for $\eta = 1$, $k_f = 0.01$, and $\sigma = 0.16$, which in dimensional terms corresponds to two cell widths. For reference, we plot the evolution for the case when growth is homogeneous. 

```{r plot remodelling foundation kf 0p01 nu 1 homog growth, echo=FALSE, fig.align="center", warning=FALSE}
directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/"
values = "Eb_0_nu_1_kf_0p01_L0_0p125_homoggrowth"

# Load the solutions
solsList <- readMat(paste(directory, "sols_", values, ".mat", sep=""))
# Load the times
timesList <- readMat(paste(directory, "times_", values, ".mat", sep=""))

numSols <- length(solsList[[1]])
solData <- NULL # Create data frame for solutions
solPointData <- NULL

# Some parameters
solCount <- 1

# # Plot solutions
for ( i in seq(2, numSols, solCount) )
{
  # Extract the solution, growth and time
  sol <- solsList[[1]][[i]][[1]]
  currentTime = timesList[[1]][[i]]
  
  # Update the solutions
  solDataFrame <- data.frame(S=sol[2,], x=sol[3,], y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,])
  solDataFrame$time = as.factor(currentTime)
  solData <- rbind(solData, solDataFrame)
}

solPlot <- ggplot()
solPlot <- solPlot + geom_path(data=solData, aes(x, -y, color=time), size=0.75)  +  ylab(TeX('$y$')) +   xlab(TeX('$x$')) + 
  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) + 
  scale_x_continuous(expand = c(0, 0), limits=c(-0.55, 1.55)) + 
  scale_y_continuous(expand = c(0, 0), limits=c(-0.65, 0.3)) +
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount))

solPlot
```

<!-- ### 'Proliferative zone' -->
<!-- \begin{align} -->
<!-- \frac{\dot{\gamma}}{\gamma} = \begin{cases} 1, & \left|s - \frac{l}{2}\right| < \sigma; \\ 0, & \mbox{otherwise.}\end{cases} -->
<!-- \end{align} -->

<!-- We parametrise the proliferative, stem-like compartment with respect  to the current arclength parameter $s$. This is analogous to tip-based growth laws used in other models. -->

<!-- ```{r plot remodelling foundation kf 0p01 nu 1 compartment-based growth, echo=FALSE, fig.align="center", warning=FALSE} -->
<!-- directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/" -->
<!-- values = "Eb_0_nu_1_kf_0p01_L0_0p125_compartmentgrowth_current_sigma_2w" -->

<!-- # Load the solutions -->
<!-- solsList <- readMat(paste(directory, "sols_", values, ".mat", sep="")) -->
<!-- gammasList <- readMat(paste(directory, "gamma_", values, ".mat", sep="")) -->
<!-- # Load the times -->
<!-- timesList <- readMat(paste(directory, "times_", values, ".mat", sep="")) -->
<!-- numSols <- length(solsList[[1]]) -->
<!-- solData <- NULL # Create data frame for solutions -->
<!-- gammaData <- NULL -->
<!-- solPointData <- NULL -->

<!-- # Some parameters -->
<!-- solCount <- 3 -->

<!-- # # Plot solutions -->
<!-- for ( i in seq(2, numSols, solCount) ) -->
<!-- { -->
<!--   # Extract the solution, growth and time -->
<!--   sol <- solsList[[1]][[i]][[1]] -->
<!--   gammaProfile <- gammasList[[1]][[i]][[1]] -->
<!--   currentTime = timesList[[1]][[i]] -->

<!--   # Update the solutions -->
<!--   solDataFrame <- data.frame(S=sol[2,], x=sol[3,], y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,]) -->
<!--   solDataFrame$time = as.factor(currentTime) -->
<!--   solData <- rbind(solData, solDataFrame) -->

<!--   gammaDataFrame <- data.frame(S=gammaProfile[1,], gamma=gammaProfile[2,]) -->
<!--   gammaDataFrame$time = as.factor(currentTime) -->
<!--   gammaData <- rbind(gammaData, gammaDataFrame) -->
<!-- } -->

<!-- solPlot <- ggplot() -->
<!-- solPlot <- solPlot + geom_path(data=solData, aes(x, -y, color=time), size=0.75)  +  ylab(TeX('$y$')) +   xlab(TeX('$x$')) +  -->
<!--   theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(-0.55, 1.55)) +  -->
<!--   scale_y_continuous(expand = c(0, 0), limits=c(-1.55, 0.05)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) -->

<!-- gammaPlot <- ggplot() -->
<!-- gammaPlot <- gammaPlot + geom_path(data=gammaData, aes(S, gamma, color=time), size=0.75)  +  ylab(TeX('$\\gamma$')) +   xlab(TeX('$S_0$')) +  -->
<!--   theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(-0.05, 1.05)) + -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) -->

<!-- ggarrange(solPlot, gammaPlot) -->

<!-- gammaPlot <- gammaPlot + scale_x_continuous(expand = c(0, 0), limits=c(0.25, 0.75)) + scale_y_continuous(expand = c(0, 0), limits=c(0.95, 100)) -->
<!-- gammaPlot -->
<!-- ``` -->

<!-- As proof of concept, we show that when growth is parametrised by the initial arclength parameter $S_0$. In this case, the 'proliferative compartment' does not change in size as time increases, leading to a shorter invagination. -->

<!-- ```{r plot remodelling foundation kf 0p01 nu 1 compartment-based growth S0, echo=FALSE, fig.align="center", warning=FALSE} -->
<!-- directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/" -->
<!-- values = "Eb_0_nu_1_kf_0p01_L0_0p125_compartmentgrowth_init_sigma_2w" -->

<!-- # Load the solutions -->
<!-- solsList <- readMat(paste(directory, "sols_", values, ".mat", sep="")) -->
<!-- gammasList <- readMat(paste(directory, "gamma_", values, ".mat", sep="")) -->
<!-- # Load the times -->
<!-- timesList <- readMat(paste(directory, "times_", values, ".mat", sep="")) -->

<!-- numSols <- length(solsList[[1]]) -->
<!-- solData <- NULL # Create data frame for solutions -->
<!-- gammaData <- NULL -->

<!-- # Some parameters -->
<!-- solCount <- 2 -->

<!-- # # Plot solutions -->
<!-- for ( i in seq(2, numSols, solCount) ) -->
<!-- { -->
<!--   # Extract the solution, growth and time -->
<!--   sol <- solsList[[1]][[i]][[1]] -->
<!--   gammaProfile <- gammasList[[1]][[i]][[1]] -->
<!--   currentTime = timesList[[1]][[i]] -->

<!--   # Update the solutions -->
<!--   solDataFrame <- data.frame(S=sol[2,], x=sol[3,], y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,]) -->
<!--   solDataFrame$time = as.factor(currentTime) -->
<!--   solData <- rbind(solData, solDataFrame) -->

<!--   gammaDataFrame <- data.frame(S=gammaProfile[1,], gamma=gammaProfile[2,]) -->
<!--   gammaDataFrame$time = as.factor(currentTime) -->
<!--   gammaData <- rbind(gammaData, gammaDataFrame) -->
<!-- } -->

<!-- solPlot <- ggplot() -->
<!-- solPlot <- solPlot + geom_path(data=solData, aes(x, -y, color=time), size=0.75)  +  ylab(TeX('$y$')) +   xlab(TeX('$x$')) +  -->
<!--   theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(-0.55, 1.55)) +  -->
<!--   scale_y_continuous(expand = c(0, 0), limits=c(-0.65, 0.15)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) -->

<!-- gammaPlot <- ggplot() -->
<!-- gammaPlot <- gammaPlot + geom_path(data=gammaData, aes(S, gamma, color=time), size=0.75)  +  ylab(TeX('$\\gamma$')) +   xlab(TeX('$S_0$')) +  -->
<!--   theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(-0.05, 1.05)) + -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) -->

<!-- ggarrange(solPlot, gammaPlot) -->

<!-- gammaPlot <- gammaPlot + scale_x_continuous(expand = c(0, 0), limits=c(0.25, 0.75)) + scale_y_continuous(expand = c(0, 0), limits=c(0.95, 5.05)) -->
<!-- gammaPlot -->
<!-- ``` -->

### Chemical signalling (Wnt)
Perhaps the fundamental driver of proliferation and hence tissue growth in the intestinal epithelium is Wnt signalling. 

\begin{align}
\frac{\dot{\gamma}}{\gamma} = W,
\end{align}
where $W$ denotes the non-dimensional profile of Wnt signalling, assumed to be in steady-state, and may be parametrised by either the inital arc length $S_0$ or the current arc length $s$:

\begin{align}
W(s) = \mathrm{e}^{-\left(\frac{s - 0.5l}{\sigma}\right)^2},
\end{align}
or
\begin{align}
W(S_0) = \mathrm{e}^{-\left(\frac{S_0 - 0.5}{\sigma}\right)^2}.
\end{align}

The below plot shows the resultant shape evolutions when Wnt is parametrised by $S_0$ and $s$, respectively, up to $t = 2.0$ days. When $W$ is parametrised with respect to the initial configuration, initial growth is much more exaggerated than when $W = W(s)$. This arises from the greater increase in net growth when $W(S_0)$:

```{r plot remodelling foundation kf 0p01 nu 1 chemical-based growth, echo=FALSE, fig.align="center", warning=FALSE}
directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/"
valuesCurrent = "Eb_0_nu_1_kf_0p01_L0_0p125_current_sigma_2w"
valuesInit = "Eb_0_nu_1_kf_0p01_L0_0p125_sigma_init_2w"

timeSamples = c(0.1, 1.0, 2.0)
solCount <- 1

# Load the solutions
solsListInit <- readMat(paste(directory, "sols_", valuesInit, ".mat", sep=""))
gammasListInit <- readMat(paste(directory, "gamma_", valuesInit, ".mat", sep=""))
# Load the times
timesListInit <- readMat(paste(directory, "times_", valuesInit, ".mat", sep=""))

numSolsInit <- length(solsListInit[[1]])
solDataInit <- NULL # Create data frame for solutions

# # Plot solutions
for ( i in seq(1, length(timeSamples), 1) )
{
  
  time = timeSamples[i]
  index = match(time, timesListInit$times)
  
  # Extract the solution, growth and time
  sol <- solsListInit[[1]][[index]][[1]]
  gammaProfile <- gammasListInit[[1]][[index]][[1]]
  currentTime = timesListInit[[1]][[index]]
  
  # Update the solutions
  solDataFrame <- data.frame(S=sol[2,], x=sol[3,], y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,])
  solDataFrame$time = as.factor(currentTime)
  solDataInit <- rbind(solDataInit, solDataFrame)

}

# Load the solutions
solsListCurrent <- readMat(paste(directory, "sols_", valuesCurrent, ".mat", sep=""))
gammasListCurrent <- readMat(paste(directory, "gamma_", valuesCurrent, ".mat", sep=""))
# Load the times
timesListCurrent <- readMat(paste(directory, "times_", valuesCurrent, ".mat", sep=""))

numSolsCurrent <- length(solsListCurrent[[1]])
solDataCurrent <- NULL # Create data frame for solutions
gammaDataCurrent <- NULL

# # Plot solutions
for ( i in seq(1, length(timeSamples), 1) )
{

  time = timeSamples[i]
  index = match(time, timesListCurrent$times)
  
  # Extract the solution, growth and time
  sol <- solsListCurrent[[1]][[index]][[1]]
  gammaProfile <- gammasListCurrent[[1]][[index]][[1]]
  currentTime = timesListCurrent[[1]][[index]]
  
  # Update the solutions
  solDataFrame <- data.frame(S=sol[2,], x=sol[3,], y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,])
  solDataFrame$time = as.factor(currentTime)
  solDataCurrent <- rbind(solDataCurrent, solDataFrame)
}

solPlot <- ggplot()
solPlot <- solPlot + geom_path(data=solDataInit, aes(x, -y, color=time), size=0.75)  +  
  geom_path(data=solDataCurrent, aes(x + 1.5, -y, color=time), size=0.75) + 
  ylab(TeX('$y$')) +   xlab(TeX('$x$')) +
  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
  scale_x_continuous(expand = c(0, 0), limits=c(-0.55, 3.05)) +
  scale_y_continuous(expand = c(0, 0), limits=c(-0.65, 0.1)) +
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(index/solCount))

solPlot

```

### Varying the width of the Wnt gradient
In the current configuration, we vary the width of the Wnt gradient. The values of $\sigma$ correspond to $\sigma = 0.16, 0.24, 0.32, 0.4$---roughly 2, 3, 4, and 5 cell widths respectively.
Here, all solutions and $\gamma$ profiles are plotted at $t = 2.0$:

```{r plot remodelling foundation kf 0p01 nu 1 chemical-based growth s compare sigma, echo=FALSE, fig.align="center", warning=FALSE}
directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/"

timeSample = 2.0

sigmaValues <- c("2w", "3w", "4w", "5w")

maxValues = "Eb_0_nu_1_kf_0p01_L0_0p125_current_sigma_2w"
maxTimesList  <- readMat(paste(directory, "times_", maxValues, ".mat", sep=""))
maxNumSols <- length(maxTimesList$times)
maxTime <- maxTimesList$times[maxNumSols]

solData <- NULL # Create data frame for solutions
gammaData <- NULL # Create data frame for gamma profiles

solPlot <- ggplot()
stressPlot <- ggplot()

for ( i in seq(1, length(sigmaValues), 1) )
{
  sigma = sigmaValues[i]
  values = paste("Eb_0_nu_1_kf_0p01_L0_0p125_current_sigma_", sigma, sep="")
  
  # Load the solutions
  solsList <- readMat(paste(directory, "sols_", values, ".mat", sep=""))
  gammasList <- readMat(paste(directory, "gamma_", values, ".mat", sep=""))
  # Load the times
  timesList <- readMat(paste(directory, "times_", values, ".mat", sep=""))
  
  # Get the solution from a sampled timepoint
  startIndex <- match(0.1, timesList$times)
  solCount <- match(0.2, timesList$times) - match(0.1, timesList$times)
  finalIndex <- match(timeSample, timesList$times)
  
  solData <- NULL
  stressData <- NULL
  for ( j in seq(startIndex, finalIndex, solCount) )
  {
    # Extract the solution, growth and time
    sol <- solsList[[1]][[j]][[1]]
    currentTime <- timesList$times[j]/maxTime
    
    # Update the solutions
    solDataFrame <- data.frame(S=sol[2,], x=sol[3,] + (i - 1)*1.5, y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,])
    solDataFrame$time = as.factor(currentTime)
    solData <- rbind(solData, solDataFrame)
    
    stressDataFrame <-data.frame(S=sol[2,] + (i - 1)*1.5, n3=sol[5,]*cos(sol[7,])+sol[6,]*sin(sol[7,]))
    stressDataFrame$time = as.factor(currentTime)
    stressData <- rbind(stressData, stressDataFrame)
    
  }
  
  solPlot <- solPlot + geom_path(data=solData, aes(x, -y, color=time), size=0.75)
  
  stressPlot <- stressPlot + geom_path(data=stressData, aes(S, n3, color=time), size=0.75)

}

# tikz(file = "PlotWntBasedCurrentGrowthFixedTimeTwoDaysVaryingSigma.tex", standAlone = T, width = 5, height = 3)

(solPlot <- solPlot + ylab('$y$') +   xlab('$x$') +
  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
  scale_x_continuous(expand = c(0, 0), limits=c(-0.05, 6.05)) +
  scale_y_continuous(expand = c(0, 0), limits=c(-0.65, 0.15)) + 
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(20)) )

(stressPlot <- stressPlot + ylab('$n_3$') +   xlab('$S_0$') +
  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
  scale_x_continuous(expand = c(0, 0), limits=c(-0.05, 6.05)) +
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(20)) )

# print(solPlot)
# dev.off()

# tools::texi2dvi('PlotWntBasedCurrentGrowthFixedTimeTwoDaysVaryingSigma.tex', pdf = T, clean = T)

```

```{r plot remodelling foundation kf 0p01 nu 1 chemical-based growth s n3 functions, echo=FALSE, fig.align="center", warning=FALSE}
directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/"

timeSample = 2.0

sigmaValues <- c("2w", "3w", "4w", "5w")

maxValues = "Eb_0_nu_1_kf_0p01_L0_0p125_current_sigma_2w"
maxTimesList  <- readMat(paste(directory, "times_", maxValues, ".mat", sep=""))
maxNumSols <- length(maxTimesList$times)
maxTime <- maxTimesList$times[maxNumSols]

stressLinearPlot <- ggplot()
stressSaturatingPlot <- ggplot()
stressBiphasicPlot <- ggplot()


for ( i in seq(1, length(sigmaValues), 1) )
{
  sigma = sigmaValues[i]
  values = paste("Eb_0_nu_1_kf_0p01_L0_0p125_current_sigma_", sigma, sep="")
  
  # Load the solutions
  solsList <- readMat(paste(directory, "sols_", values, ".mat", sep=""))
  gammasList <- readMat(paste(directory, "gamma_", values, ".mat", sep=""))
  # Load the times
  timesList <- readMat(paste(directory, "times_", values, ".mat", sep=""))
  
  # Get the solution from a sampled timepoint
  startIndex <- match(0.1, timesList$times)
  solCount <- match(0.2, timesList$times) - match(0.1, timesList$times)
  finalIndex <- match(timeSample, timesList$times)
  
  stressDataLinear <- NULL
  stressDataSaturating <- NULL
  stressDataBiphasic <- NULL
  
  for ( j in seq(startIndex, finalIndex, solCount) )
  {
    # Extract the solution, growth and time
    sol <- solsList[[1]][[j]][[1]]
    currentTime <- timesList$times[j]/maxTime
    
    # Update the solutions
    stressDataLinearFrame <-data.frame(S=sol[2,] + (i - 1)*1.5, f=sol[5,]*cos(sol[7,])+sol[6,]*sin(sol[7,]))
    stressDataLinearFrame$time = as.factor(currentTime)
    stressDataLinear <- rbind(stressDataLinear, stressDataLinearFrame)
    
    stressDataSaturatingFrame <-data.frame(S=sol[2,] + (i - 1)*1.5, f=tanh(sol[5,]*cos(sol[7,])+sol[6,]*sin(sol[7,])))
    stressDataSaturatingFrame$time = as.factor(currentTime)
    stressDataSaturating <- rbind(stressDataSaturating, stressDataSaturatingFrame)
    
    
    stressDataBiphasicFrame <-data.frame(S=sol[2,] + (i - 1)*1.5, f=sol[5,]*cos(sol[7,])+sol[6,]*sin(sol[7,])/(1 + (sol[5,]*cos(sol[7,])+sol[6,]*sin(sol[7,]))^2))
    stressDataBiphasicFrame$time = as.factor(currentTime)
    stressDataBiphasic <- rbind(stressDataBiphasic, stressDataBiphasicFrame)
    
  }
  
  
  stressLinearPlot <- stressPlot + geom_path(data=stressDataLinear, aes(S, f, color=time), size=0.75)
  stressSaturatingPlot <- stressPlot + geom_path(data=stressDataSaturating, aes(S, f, color=time), size=0.75)
  stressBiphasicPlot <- stressPlot + geom_path(data=stressDataBiphasic, aes(S, f, color=time), size=0.75)


}

# tikz(file = "PlotWntBasedCurrentGrowthFixedTimeTwoDaysVaryingSigma.tex", standAlone = T, width = 5, height = 3)

(stressLinearPlot <- stressLinearPlot +
  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
  scale_x_continuous(expand = c(0, 0), limits=c(-0.05, 6.05)) +
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(20)) )

(stressSaturatingPlot <- stressSaturatingPlot +
  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
  scale_x_continuous(expand = c(0, 0), limits=c(-0.05, 6.05)) +
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(20)) )

(stressBiphasicPlot <- stressBiphasicPlot +
  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
  scale_x_continuous(expand = c(0, 0), limits=c(-0.05, 6.05)) +
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(20)) )

# print(solPlot)
# dev.off()

# tools::texi2dvi('PlotWntBasedCurrentGrowthFixedTimeTwoDaysVaryingSigma.tex', pdf = T, clean = T)

```

```{r plot remodelling foundation kf 0p01 nu 1 chemical-based growth s n1 functions, echo=FALSE, fig.align="center", warning=FALSE}
directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/"

timeSample = 2.0

sigmaValues <- c("2w", "3w", "4w", "5w")

maxValues = "Eb_0_nu_1_kf_0p01_L0_0p125_current_sigma_2w"
maxTimesList  <- readMat(paste(directory, "times_", maxValues, ".mat", sep=""))
maxNumSols <- length(maxTimesList$times)
maxTime <- maxTimesList$times[maxNumSols]

n1LinearPlot <- ggplot()
n1SaturatingPlot <- ggplot()
n1BiphasicPlot <- ggplot()


for ( i in seq(1, length(sigmaValues), 1) )
{
  sigma = sigmaValues[i]
  values = paste("Eb_0_nu_1_kf_0p01_L0_0p125_current_sigma_", sigma, sep="")
  
  # Load the solutions
  solsList <- readMat(paste(directory, "sols_", values, ".mat", sep=""))
  gammasList <- readMat(paste(directory, "gamma_", values, ".mat", sep=""))
  # Load the times
  timesList <- readMat(paste(directory, "times_", values, ".mat", sep=""))
  
  # Get the solution from a sampled timepoint
  startIndex <- match(0.1, timesList$times)
  solCount <- match(0.2, timesList$times) - match(0.1, timesList$times)
  finalIndex <- match(timeSample, timesList$times)
  
  stressDataLinear <- NULL
  stressDataSaturating <- NULL
  stressDataBiphasic <- NULL
  
  for ( j in seq(startIndex, finalIndex, solCount) )
  {
    # Extract the solution, growth and time
    sol <- solsList[[1]][[j]][[1]]
    currentTime <- timesList$times[j]/maxTime
    
    # Update the solutions
    stressDataLinearFrame <-data.frame(S=sol[2,] + (i - 1)*1.5, f= -sol[5,]*sin(sol[7,])+sol[6,]*cos(sol[7,]))
    stressDataLinearFrame$time = as.factor(currentTime)
    stressDataLinear <- rbind(stressDataLinear, stressDataLinearFrame)
    
    stressDataSaturatingFrame <-data.frame(S=sol[2,] + (i - 1)*1.5, f=tanh(-sol[5,]*sin(sol[7,])+sol[6,]*cos(sol[7,])))
    stressDataSaturatingFrame$time = as.factor(currentTime)
    stressDataSaturating <- rbind(stressDataSaturating, stressDataSaturatingFrame)
    
    
    stressDataBiphasicFrame <-data.frame(S=sol[2,] + (i - 1)*1.5, f=-sol[5,]*sin(sol[7,])+sol[6,]*cos(sol[7,])/(1 + (-sol[5,]*sin(sol[7,])+sol[6,]*cos(sol[7,]))^2))
    stressDataBiphasicFrame$time = as.factor(currentTime)
    stressDataBiphasic <- rbind(stressDataBiphasic, stressDataBiphasicFrame)
    
  }
  
  
  n1LinearPlot <- stressPlot + geom_path(data=stressDataLinear, aes(S, f, color=time), size=0.75)
  n1SaturatingPlot <- stressPlot + geom_path(data=stressDataSaturating, aes(S, f, color=time), size=0.75)
  n1BiphasicPlot <- stressPlot + geom_path(data=stressDataBiphasic, aes(S, f, color=time), size=0.75)


}

# tikz(file = "PlotWntBasedCurrentGrowthFixedTimeTwoDaysVaryingSigma.tex", standAlone = T, width = 5, height = 3)

(n1LinearPlot <- n1LinearPlot +
  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
  scale_x_continuous(expand = c(0, 0), limits=c(-0.05, 6.05)) +
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(20)) )

(n1SaturatingPlot <- n1SaturatingPlot +
  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
  scale_x_continuous(expand = c(0, 0), limits=c(-0.05, 6.05)) +
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(20)) )

(n1BiphasicPlot <- n1BiphasicPlot +
  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
  scale_x_continuous(expand = c(0, 0), limits=c(-0.05, 6.05)) +
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(20)) )

# print(solPlot)
# dev.off()

# tools::texi2dvi('PlotWntBasedCurrentGrowthFixedTimeTwoDaysVaryingSigma.tex', pdf = T, clean = T)

```

<!-- However, the final profiles look very different: -->
<!-- ```{r plot remodelling foundation kf 0p01 nu 1 chemical-based growth s compare sigma and final shape, echo=FALSE, fig.align="center", warning=FALSE} -->
<!-- directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/" -->

<!-- timeSample = 1.0 -->

<!-- sigmaValues <- c("2w", "3w", "4w", "5w") -->

<!-- maxValues = "Eb_0_nu_1_kf_0p01_L0_0p125_current_sigma_2w" -->
<!-- maxTimesList  <- readMat(paste(directory, "times_", maxValues, ".mat", sep="")) -->
<!-- maxNumSols <- length(maxTimesList$times) -->
<!-- maxTime <- maxTimesList$times[maxNumSols] -->

<!-- solData <- NULL # Create data frame for solutions -->
<!-- gammaData <- NULL # Create data frame for gamma profiles -->

<!-- solPlot <- ggplot() -->

<!-- for ( i in seq(1, length(sigmaValues), 1) ) -->
<!-- { -->
<!--   sigma = sigmaValues[i] -->
<!--   values = paste("Eb_0_nu_1_kf_0p01_L0_0p125_current_sigma_", sigma, sep="") -->

<!--   # Load the solutions -->
<!--   solsList <- readMat(paste(directory, "sols_", values, ".mat", sep="")) -->
<!--   gammasList <- readMat(paste(directory, "gamma_", values, ".mat", sep="")) -->
<!--   # Load the times -->
<!--   timesList <- readMat(paste(directory, "times_", values, ".mat", sep="")) -->

<!--   # Get the solution from a sampled timepoint -->
<!--   finalIndex <- length(timesList$times) -->
<!--   startIndex <- 2 -->
<!--   if (i == 1) -->
<!--   { -->
<!--     solCount <- 5 -->
<!--   } -->
<!--   else -->
<!--   { -->
<!--     solCount <- 5 -->
<!--   } -->

<!--   for ( j in seq(startIndex, finalIndex, solCount) ) -->
<!--   { -->
<!--     # Extract the solution, growth and time -->
<!--     sol <- solsList[[1]][[j]][[1]] -->
<!--     currentTime <- timesList$times[j]/maxTime -->

<!--     # Update the solutions -->
<!--     solDataFrame <- data.frame(S=sol[2,], x=sol[3,] + (i - 1)*1.5, y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,]) -->
<!--     solDataFrame$time = as.factor(currentTime) -->
<!--     solData <- rbind(solData, solDataFrame) -->

<!--   } -->

<!--   solPlot <- solPlot + geom_path(data=solData, aes(x, -y, color=time), size=0.75) -->
<!-- } -->

<!-- tikz(file = "PlotWntBasedCurrentGrowthFinalTimeVaryingSigma.tex", standAlone = T, width = 5, height = 3) -->

<!-- solPlot <- solPlot + ylab('$y$') +   xlab('$x$') + -->
<!--   theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) + -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(-0.05, 6.05)) + -->
<!--   scale_y_continuous(expand = c(0, 0), limits=c(-1.65, 0.15)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(65)) -->

<!-- print(solPlot) -->
<!-- dev.off() -->

<!-- tools::texi2dvi('PlotWntBasedCurrentGrowthFinalTimeVaryingSigma.tex', pdf = T, clean = T) -->

<!-- ``` -->

<!-- ### Mechanosensitivity -->
<!-- ```{r plot remodelling foundation stresses homog, echo=FALSE, fig.align="center", warning=FALSE} -->
<!-- directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/" -->
<!-- values = "Eb_1_nu_1_kf_0p01_L0_0p125_homoggrowth" -->

<!-- # Load the solutions -->
<!-- solsList <- readMat(paste(directory, "sols_", values, ".mat", sep="")) -->
<!-- # Load the times -->
<!-- timesList <- readMat(paste(directory, "times_", values, ".mat", sep="")) -->

<!-- numSols <- length(solsList[[1]]) -->
<!-- solData <- NULL # Create data frame for solutions -->
<!-- stressData <- NULL -->

<!-- # Some parameters -->
<!-- solCount <- 2 -->

<!-- # # Plot solutions -->
<!-- for ( i in seq(2, numSols, solCount) ) -->
<!-- { -->
<!--   # Extract the solution, growth and time -->
<!--   sol <- solsList[[1]][[i]][[1]] -->
<!--   currentTime = timesList[[1]][[i]] -->

<!--   # Update the solutions -->
<!--   solDataFrame <- data.frame(S=sol[2,], x=sol[3,], y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,]) -->
<!--   solDataFrame$time = toString(currentTime) -->
<!--   solData <- rbind(solData, solDataFrame) -->

<!--   stressDataFrame <- data.frame(S=sol[1,], nX=sol[5,], nY=sol[6,], n1=sol[6,]*cos(sol[7,])-sol[5,]*sin(sol[7,]), n3=sol[6,]*sin(sol[7,])+sol[5,]*cos(sol[7,]),sigma=sqrt((sol[3,] - sol[2,])^2 + (sol[4,])^2)) -->
<!--   stressDataFrame$time = toString(currentTime) -->
<!--   stressData <- rbind(stressData, stressDataFrame) -->
<!-- } -->

<!-- solPlotnX <- ggplot() +  geom_path(data=solData, aes(S, nX, color=time), size=0.75)  +  ylab(TeX('$n_x$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- solPlotnY <- ggplot() +  geom_path(data=solData, aes(S, -nY, color=time), size=0.75)  +  ylab(TeX('$n_y$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- solPlotn1 <- ggplot() +  geom_path(data=stressData, aes(S, n1, color=time), size=0.75)  +  ylab(TeX('$n_1$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- solPlotn3 <- ggplot() +  geom_path(data=stressData, aes(S, n3, color=time), size=0.75)  +  ylab(TeX('$n_3$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- solPlotnDelta <-  ggplot() +  geom_path(data=stressData, aes(S, sqrt((n1)^2 + (n3)^2), color=time), size=0.75)  +  ylab(TeX('$\\sqrt{(n_3)^2 + (n_1)^2}$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- ggarrange(solPlotnX, solPlotnY, solPlotn1, solPlotn3) -->
<!-- ``` -->

<!-- ```{r plot remodelling foundation stresses sigma 2w, echo=FALSE, fig.align="center", warning=FALSE} -->
<!-- directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/" -->
<!-- values = "Eb_0_nu_1_kf_0p01_L0_0p125_current_sigma_2w" -->

<!-- # Load the solutions -->
<!-- solsList <- readMat(paste(directory, "sols_", values, ".mat", sep="")) -->
<!-- # Load the times -->
<!-- timesList <- readMat(paste(directory, "times_", values, ".mat", sep="")) -->

<!-- numSols <- length(solsList[[1]]) -->
<!-- solData <- NULL # Create data frame for solutions -->
<!-- stressData <- NULL -->

<!-- # Some parameters -->
<!-- solCount <- 2 -->

<!-- # # Plot solutions -->
<!-- for ( i in seq(2, numSols, solCount) ) -->
<!-- { -->
<!--   # Extract the solution, growth and time -->
<!--   sol <- solsList[[1]][[i]][[1]] -->
<!--   currentTime = timesList[[1]][[i]] -->

<!--   # Update the solutions -->
<!--   solDataFrame <- data.frame(S=sol[2,], x=sol[3,], y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,]) -->
<!--   solDataFrame$time = toString(currentTime) -->
<!--   solData <- rbind(solData, solDataFrame) -->

<!--   stressDataFrame <- data.frame(S=sol[1,], nX=sol[5,], nY=sol[6,], n1=sol[6,]*cos(sol[7,])-sol[5,]*sin(sol[7,]), n3=sol[6,]*sin(sol[7,])+sol[5,]*cos(sol[7,]),sigma=sqrt((sol[3,] - sol[2,])^2 + (sol[4,])^2)) -->
<!--   stressDataFrame$time = toString(currentTime) -->
<!--   stressData <- rbind(stressData, stressDataFrame) -->
<!-- } -->

<!-- solPlotnX <- ggplot() +  geom_path(data=solData, aes(S, nX, color=time), size=0.75)  +  ylab(TeX('$n_x$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- solPlotnY <- ggplot() +  geom_path(data=solData, aes(S, -nY, color=time), size=0.75)  +  ylab(TeX('$n_y$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- solPlotn1 <- ggplot() +  geom_path(data=stressData, aes(S, n1, color=time), size=0.75)  +  ylab(TeX('$n_1$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- solPlotn3 <- ggplot() +  geom_path(data=stressData, aes(S, n3, color=time), size=0.75)  +  ylab(TeX('$n_3$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- ggarrange(solPlotnX, solPlotnY, solPlotn1, solPlotn3) -->
<!-- ``` -->

<!-- ```{r plot remodelling foundation stresses sigma 3w, echo=FALSE, fig.align="center", warning=FALSE} -->
<!-- directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/" -->
<!-- values = "Eb_0_nu_1_kf_0p01_L0_0p125_current_sigma_3w" -->

<!-- # Load the solutions -->
<!-- solsList <- readMat(paste(directory, "sols_", values, ".mat", sep="")) -->
<!-- # Load the times -->
<!-- timesList <- readMat(paste(directory, "times_", values, ".mat", sep="")) -->

<!-- numSols <- length(solsList[[1]]) -->
<!-- solData <- NULL # Create data frame for solutions -->
<!-- stressData <- NULL -->

<!-- # Some parameters -->
<!-- solCount <- 2 -->

<!-- # # Plot solutions -->
<!-- for ( i in seq(2, numSols, solCount) ) -->
<!-- { -->
<!--   # Extract the solution, growth and time -->
<!--   sol <- solsList[[1]][[i]][[1]] -->
<!--   currentTime = timesList[[1]][[i]] -->

<!--   # Update the solutions -->
<!--   solDataFrame <- data.frame(S=sol[2,], x=sol[3,], y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,]) -->
<!--   solDataFrame$time = toString(currentTime) -->
<!--   solData <- rbind(solData, solDataFrame) -->

<!--   stressDataFrame <- data.frame(S=sol[1,], nX=sol[5,], nY=sol[6,], n1=sol[6,]*cos(sol[7,])-sol[5,]*sin(sol[7,]), n3=sol[6,]*sin(sol[7,])+sol[5,]*cos(sol[7,]),sigma=sqrt((sol[3,] - sol[2,])^2 + (sol[4,])^2)) -->
<!--   stressDataFrame$time = toString(currentTime) -->
<!--   stressData <- rbind(stressData, stressDataFrame) -->
<!-- } -->

<!-- solPlotnX <- ggplot() +  geom_path(data=solData, aes(S, nX, color=time), size=0.75)  +  ylab(TeX('$n_x$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- solPlotnY <- ggplot() +  geom_path(data=solData, aes(S, -nY, color=time), size=0.75)  +  ylab(TeX('$n_y$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- solPlotn1 <- ggplot() +  geom_path(data=stressData, aes(S, n1, color=time), size=0.75)  +  ylab(TeX('$n_1$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- solPlotn3 <- ggplot() +  geom_path(data=stressData, aes(S, n3, color=time), size=0.75)  +  ylab(TeX('$n_3$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- ggarrange(solPlotnX, solPlotnY, solPlotn1, solPlotn3) -->
<!-- ``` -->

<!-- ```{r plot remodelling foundation stresses sigma 4w, echo=FALSE, fig.align="center", warning=FALSE} -->
<!-- directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/" -->
<!-- values = "Eb_0_nu_1_kf_0p01_L0_0p125_current_sigma_4w" -->

<!-- # Load the solutions -->
<!-- solsList <- readMat(paste(directory, "sols_", values, ".mat", sep="")) -->
<!-- # Load the times -->
<!-- timesList <- readMat(paste(directory, "times_", values, ".mat", sep="")) -->

<!-- numSols <- length(solsList[[1]]) -->
<!-- solData <- NULL # Create data frame for solutions -->
<!-- stressData <- NULL -->

<!-- # Some parameters -->
<!-- solCount <- 2 -->

<!-- # # Plot solutions -->
<!-- for ( i in seq(2, numSols, solCount) ) -->
<!-- { -->
<!--   # Extract the solution, growth and time -->
<!--   sol <- solsList[[1]][[i]][[1]] -->
<!--   currentTime = timesList[[1]][[i]] -->

<!--   # Update the solutions -->
<!--   solDataFrame <- data.frame(S=sol[2,], x=sol[3,], y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,]) -->
<!--   solDataFrame$time = toString(currentTime) -->
<!--   solData <- rbind(solData, solDataFrame) -->

<!--   stressDataFrame <- data.frame(S=sol[1,], nX=sol[5,], nY=sol[6,], n1=sol[6,]*cos(sol[7,])-sol[5,]*sin(sol[7,]), n3=sol[6,]*sin(sol[7,])+sol[5,]*cos(sol[7,]),sigma=sqrt((sol[3,] - sol[2,])^2 + (sol[4,])^2)) -->
<!--   stressDataFrame$time = toString(currentTime) -->
<!--   stressData <- rbind(stressData, stressDataFrame) -->
<!-- } -->

<!-- solPlotnX <- ggplot() +  geom_path(data=solData, aes(S, nX, color=time), size=0.75)  +  ylab(TeX('$n_x$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- solPlotnY <- ggplot() +  geom_path(data=solData, aes(S, -nY, color=time), size=0.75)  +  ylab(TeX('$n_y$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- solPlotn1 <- ggplot() +  geom_path(data=stressData, aes(S, n1, color=time), size=0.75)  +  ylab(TeX('$n_1$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- solPlotn3 <- ggplot() +  geom_path(data=stressData, aes(S, n3, color=time), size=0.75)  +  ylab(TeX('$n_3$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +  -->
<!--   scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) +  -->
<!--   scale_x_continuous(expand = c(0, 0), limits=c(0,1.05)) -->

<!-- ggarrange(solPlotnX, solPlotnY, solPlotn1, solPlotn3) -->
<!-- ``` -->

<!-- Having first looked at growth in the absence of mechanical coupling, we begin to investigate the effects of mechanical feedback laws. -->

<!-- In the absence of mechanical feedback, Wnt parametrised in the current configuration produced the longest crypt. As such, we will explore mechanosensitivity with respect to Wnt-dependent growth. -->

<!-- After exploration of the foundation laws, a suggested law to account for mechanical inhibition/promotion of epithelial-to-epithelial stress is -->
<!-- \begin{align} -->
<!-- \frac{\dot{\gamma}}{\gamma} = W(s) + \mu_3(n_3 - n_3*),\qquad\mbox{where}\qquad n_3 = n_x\cos\theta + n_y\sin\theta. -->
<!-- \end{align} -->

<!-- We set $n_3^* = 0$ and explore the effect of $\mu_3 > 0$ that increases in magnitude.    -->