---
title: "Modelling cellular mechanical properties in the crypt"
author: "Axel Almet"
date: "2 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(cowplot)
library(R.matlab)
library(latex2exp)
library(RColorBrewer)
library(ggpubr)
library(plotly)
```

# Generalised-set up
We model the crypt as a planar, growing, inextensible elastic rod attached to an underlying foundation, whose attachments remodel over time to the current rod shape. Tissue growth is encapsulated by the growth stretch $\gamma$, modelling a local increase of the initial arc length $S_0$; that is, 
\begin{align}
\gamma = \frac{\partial s}{\partial S_0}.
\end{align}
The equations describing geometric constraints and balance of linear and angular momentum are given by:
\begin{align}
&x' = \gamma\cos\theta,\\
&y' = \gamma\sin\theta,\\
&n_x' = k\gamma(x - p_x),\qquad\mbox{where}\qquad \dot{p}_x = \rho(x - p_x),\\
&n_y' = k\gamma(y - p_y),\qquad\mbox{where}\qquad \dot{p}_y = \rho(y - p_y),\\
&\theta' = E_b\gamma m,\\
&m' = \gamma(n_x\sin\theta - n_y\cos\theta).
\end{align}
For simplicity, we assume the rod is clamped horizontally at the boundaries:
\begin{align}
x(0) = 0, \qquad x(1) = 0, \qquad y(0) = y(1) =  y_0, \qquad \theta(0) = \theta(1) = 0.
\end{align}

# Specifying $E_b$

\begin{align}
E_b(S_0) = 1 - \frac{1}{2}e^{-\left(\frac{S_0 - 0.5}{\sigma_E}\right)^2}.
\end{align}

\begin{align}
E_b(s) = 1 - \frac{1}{2}e^{-\left(\frac{s - 0.5l}{\sigma_E}\right)^2}.
\end{align}

# Results

```{r plot remodelling foundation kf 0p01 nu 1 compare sigma init, echo=FALSE, fig.align="center", warning=FALSE}
directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/"

sigmaValues <- c("2w", "3w", "4w", "5w")
timeSample = 1.0

maxValues = "Eb_0p5_init_sigmaE_w_nu_1_kf_0p01_L0_0p125_homoggrowth"
maxTimesList  <- readMat(paste(directory, "times_", maxValues, ".mat", sep=""))
maxNumSols <- length(maxTimesList$times)
maxTime <- maxTimesList$times[maxNumSols]

solCount <- 1
solPlot <- ggplot()

count <- 1
for (i in seq(1, length(sigmaValues), 1))
{
  sigma = i
  
  values = paste("Eb_0p5_init_sigmaE_", sigmaValues[i], "_nu_1_kf_0p01_L0_0p125_homoggrowth", sep="")
  
  # Load the solutions
  solsList <- readMat(paste(directory, "sols_", values, ".mat", sep=""))
  # Load the times
  timesList <- readMat(paste(directory, "times_", values, ".mat", sep=""))
  
  solData <- NULL 
  # finalIndex = match(timeSample, timesList$times)
  finalIndex = length(timesList$times)
  
  if (i == length(sigmaValues))
  {
    startIndex <- 2
    solCount <- 2
  }
  else
  {
    startIndex <- 2
    solCount <- 2
  }
  
  for ( j in seq(startIndex, finalIndex, solCount) )
  {
    # Extract the solution, growth and time
    sol <- solsList[[1]][[j]][[1]]
    currentTime <- timesList$times[j]/maxTime
    
    # Update the solutions
    solDataFrame <- data.frame(S=sol[2,], x=sol[3,] + (i - 1)*1.5, y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,])
    solDataFrame$time = as.factor(currentTime)
    solData <- rbind(solData, solDataFrame)

  }
  
  solPlot <- solPlot + geom_path(data=solData, aes(x, - y, color=time), size=0.75)
  
}

solPlot <- solPlot +  ylab(TeX('$y$')) +   xlab(TeX('$x$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(maxNumSols/solCount)) +
  scale_x_continuous(expand = c(0, 0), limits=c(0, length(sigmaValues)*1.5))

solPlot
```

```{r plot remodelling foundation kf 0p01 nu 1 compare sigma current, echo=FALSE, fig.align="center", warning=FALSE}
directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/"

sigmaValues <- c("w", "2w", "3w", "4w", "5w")
timeSample = 1.0

maxValues = "Eb_0p5_current_sigmaE_w_nu_1_kf_0p01_L0_0p125_homoggrowth"
maxTimesList  <- readMat(paste(directory, "times_", maxValues, ".mat", sep=""))
maxNumSols <- length(maxTimesList$times)
maxTime <- maxTimesList$times[maxNumSols]

solCount <- 1
solPlot <- ggplot()

for (i in seq(1, length(sigmaValues), 1))
{
  sigma = i
  
  values = paste("Eb_0p5_current_sigmaE_", sigmaValues[i], "_nu_1_kf_0p01_L0_0p125_homoggrowth", sep="")
  
  # Load the solutions
  solsList <- readMat(paste(directory, "sols_", values, ".mat", sep=""))
  # Load the times
  timesList <- readMat(paste(directory, "times_", values, ".mat", sep=""))
  
  solData <- NULL 
  # finalIndex = match(timeSample, timesList$times)
  finalIndex = length(timesList$times)
  
  if (i == length(sigmaValues))
  {
    startIndex <- 2
    solCount <- 2
  }
  else
  {
    startIndex <- 2
    solCount <- 2
  }
  
  for ( j in seq(startIndex, finalIndex, solCount) )
  {
    # Extract the solution, growth and time
    sol <- solsList[[1]][[j]][[1]]
    currentTime <- timesList$times[j]/maxTime
    
    # Update the solutions
    solDataFrame <- data.frame(S=sol[2,], x=sol[3,] + (i - 1)*1.5, y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,])
    solDataFrame$time = as.factor(currentTime)
    solData <- rbind(solData, solDataFrame)

  }
  
  solPlot <- solPlot + geom_path(data=solData, aes(x, - y, color=time), size=0.75)
  
}

solPlot <- solPlot +  ylab(TeX('$y$')) +   xlab(TeX('$x$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(maxNumSols/solCount)) +
  scale_x_continuous(expand = c(0, 0), limits=c(0, length(sigmaValues)*1.5))

solPlot
```

```{r plot sigmaE 2w nu 1 stresses, echo=FALSE, fig.align="center", warning=FALSE}
directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/"
values = "Eb_0p5_current_sigmaE_2w_nu_1_kf_0p01_L0_0p125_homoggrowth"

# Load the solutions
solsList <- readMat(paste(directory, "sols_", values, ".mat", sep=""))
# Load the times
timesList <- readMat(paste(directory, "times_", values, ".mat", sep=""))

numSols <- length(solsList[[1]])
solData <- NULL # Create data frame for solutions
stressData <- NULL

# Some parameters
solCount <- 2

# # Plot solutions
for ( i in seq(2, numSols, solCount) )
{
  # Extract the solution, growth and time
  sol <- solsList[[1]][[i]][[1]]
  currentTime = timesList[[1]][[i]]

  # Update the solutions
  solDataFrame <- data.frame(S=sol[2,], x=sol[3,], y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,])
  solDataFrame$time = toString(currentTime)
  solData <- rbind(solData, solDataFrame)
  
  stressDataFrame <- data.frame(S=sol[1,], nX=sol[5,], nY=sol[6,], n1=sol[6,]*cos(sol[7,]) - sol[5,]*sin(sol[7,]), n3=sol[5,]*cos(sol[7,]) + sol[6,]*sin(sol[7,]))
  stressDataFrame$time = toString(currentTime)
  stressData <- rbind(stressData, stressDataFrame)
}

solPlotnX <- ggplot() +  geom_path(data=solData, aes(S, nX, color=time), size=0.75)  +  ylab(TeX('$n_x$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) + 
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) + 
            scale_x_continuous(expand = c(0, 0), limits=c(0,1.05))

solPlotnY <- ggplot() +  geom_path(data=solData, aes(S, -nY, color=time), size=0.75)  +  ylab(TeX('$n_y$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) + 
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) + 
            scale_x_continuous(expand = c(0, 0), limits=c(0,1.05))

solPlotn1 <- ggplot() +  geom_path(data=stressData, aes(S, n1, color=time), size=0.75)  +  ylab(TeX('$n_1$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) + 
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) + 
            scale_x_continuous(expand = c(0, 0), limits=c(0,1.05))

solPlotn3 <- ggplot() +  geom_path(data=stressData, aes(S, n3, color=time), size=0.75)  +  ylab(TeX('$n_3$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) + 
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) + 
            scale_x_continuous(expand = c(0, 0), limits=c(0,1.05))

solPlotsigma <- ggplot() +  geom_path(data=stressData, aes(S, sigma, color=time), size=0.75)  +  ylab(TeX('$\\sigma$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) + 
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) + 
            scale_x_continuous(expand = c(0, 0), limits=c(0,1.05))

ggarrange(solPlotnX, solPlotnY, solPlotn3, solPlotn1)

```

```{r plot remodelling foundation kf 0p01 nu 10 compare sigma current, echo=FALSE, fig.align="center", warning=FALSE}
directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/"

sigmaValues <- c("w", "2w", "3w", "4w", "5w")
timeSample = 1.0

maxValues = "Eb_0p5_current_sigmaE_w_nu_10_kf_0p01_L0_0p125_homoggrowth"
maxTimesList  <- readMat(paste(directory, "times_", maxValues, ".mat", sep=""))
maxNumSols <- length(maxTimesList$times)
maxTime <- maxTimesList$times[maxNumSols]

solCount <- 1
solPlot <- ggplot()

count <- 1
for (i in seq(1, length(sigmaValues), 1))
{
  sigma = i
  
  values = paste("Eb_0p5_current_sigmaE_", sigmaValues[i], "_nu_10_kf_0p01_L0_0p125_homoggrowth", sep="")
  
  # Load the solutions
  solsList <- readMat(paste(directory, "sols_", values, ".mat", sep=""))
  # Load the times
  timesList <- readMat(paste(directory, "times_", values, ".mat", sep=""))
  
  solData <- NULL 
  # finalIndex = match(timeSample, timesList$times)
  finalIndex = length(timesList$times)
  
  if (i == length(sigmaValues))
  {
    startIndex <- 2
    solCount <- 2
  }
  else
  {
    startIndex <- 2
    solCount <- 2
  }
  
  for ( j in seq(startIndex, finalIndex, solCount) )
  {
    # Extract the solution, growth and time
    sol <- solsList[[1]][[j]][[1]]
    currentTime <- timesList$times[j]/maxTime
    
    # Update the solutions
    solDataFrame <- data.frame(S=sol[2,], x=sol[3,] + (i - 1)*1.5, y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,])
    solDataFrame$time = as.factor(currentTime)
    solDataFrame$sigma = as.factor(sigma)
    solData <- rbind(solData, solDataFrame)

  }
  
  solPlot <- solPlot + geom_path(data=solData, aes(x, - y, color=time), size=0.75)
  
}

solPlot <- solPlot +  ylab(TeX('$y$')) +   xlab(TeX('$x$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(maxNumSols/solCount)) +
  scale_x_continuous(expand = c(0, 0), limits=c(0, length(sigmaValues)*1.5))

solPlot
```

```{r plot sigmaE current 2w nu 10 stresses, echo=FALSE, fig.align="center", warning=FALSE}
directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/"
values = "Eb_0p5_current_sigmaE_2w_nu_10_kf_0p01_L0_0p125_homoggrowth"

# Load the solutions
solsList <- readMat(paste(directory, "sols_", values, ".mat", sep=""))
# Load the times
timesList <- readMat(paste(directory, "times_", values, ".mat", sep=""))

numSols <- length(solsList[[1]])
solData <- NULL # Create data frame for solutions
stressData <- NULL

# Some parameters
solCount <- 2

# # Plot solutions
for ( i in seq(2, numSols, solCount) )
{
  # Extract the solution, growth and time
  sol <- solsList[[1]][[i]][[1]]
  currentTime = timesList[[1]][[i]]

  # Update the solutions
  solDataFrame <- data.frame(S=sol[2,], x=sol[3,], y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,])
  solDataFrame$time = toString(currentTime)
  solData <- rbind(solData, solDataFrame)
  
  stressDataFrame <- data.frame(S=sol[1,], nX=sol[5,], nY=sol[6,], n1=sol[6,]*cos(sol[7,]) - sol[5,]*sin(sol[7,]), n3=sol[5,]*cos(sol[7,]) + sol[6,]*sin(sol[7,]))
  stressDataFrame$time = toString(currentTime)
  stressData <- rbind(stressData, stressDataFrame)
}

solPlotnX <- ggplot() +  geom_path(data=solData, aes(S, nX, color=time), size=0.75)  +  ylab(TeX('$n_x$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) + 
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) + 
            scale_x_continuous(expand = c(0, 0), limits=c(0,1.05))

solPlotnY <- ggplot() +  geom_path(data=solData, aes(S, -nY, color=time), size=0.75)  +  ylab(TeX('$n_y$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) + 
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) + 
            scale_x_continuous(expand = c(0, 0), limits=c(0,1.05))

solPlotn1 <- ggplot() +  geom_path(data=stressData, aes(S, n1, color=time), size=0.75)  +  ylab(TeX('$n_1$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) + 
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) + 
            scale_x_continuous(expand = c(0, 0), limits=c(0,1.05))

solPlotn3 <- ggplot() +  geom_path(data=stressData, aes(S, n3, color=time), size=0.75)  +  ylab(TeX('$n_3$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) + 
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) + 
            scale_x_continuous(expand = c(0, 0), limits=c(0,1.05))

solPlotsigma <- ggplot() +  geom_path(data=stressData, aes(S, sigma, color=time), size=0.75)  +  ylab(TeX('$\\sigma$')) +   xlab(TeX('$S_0$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) + 
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(numSols/solCount)) + 
            scale_x_continuous(expand = c(0, 0), limits=c(0,1.05))

ggarrange(solPlotnX, solPlotnY, solPlotn3, solPlotn1)

```

```{r plot remodelling foundation kf 0p01 nu 10 compare b, echo=FALSE, fig.align="center", warning=FALSE}
directory = "~/Documents/MorphoelasticCrypt/MATLAB/Solutions/RemodellingFoundation/"

sigmaValues <- c("0p1", "0p5", "0p9")
timeSample = 1.25

maxValues = "Eb_0p9_init_sigmaE_w_nu_10_kf_0p01_L0_0p125_homoggrowth"
maxTimesList  <- readMat(paste(directory, "times_", maxValues, ".mat", sep=""))
maxNumSols <- length(maxTimesList$times)
maxTime <- maxTimesList$times[maxNumSols]

solCount <- 1
solPlot <- ggplot()

count <- 1
for (i in seq(1, length(sigmaValues), 1))
{
  sigma = i
  
  values = paste("Eb_", sigmaValues[i], "_init_sigmaE_w_nu_10_kf_0p01_L0_0p125_homoggrowth", sep="")
  
  # Load the solutions
  solsList <- readMat(paste(directory, "sols_", values, ".mat", sep=""))
  # Load the times
  timesList <- readMat(paste(directory, "times_", values, ".mat", sep=""))
  
  solData <- NULL 
  finalIndex = match(timeSample, timesList$times)
  # finalIndex = length(timesList$times)
  
  if (i == length(sigmaValues))
  {
    startIndex <- 2
    solCount <- 2
  }
  else
  {
    startIndex <- 2
    solCount <- 2
  }
  
  for ( j in seq(startIndex, finalIndex, solCount) )
  {
    # Extract the solution, growth and time
    sol <- solsList[[1]][[j]][[1]]
    currentTime <- timesList$times[j]/maxTime
    
    # Update the solutions
    solDataFrame <- data.frame(S=sol[2,], x=sol[3,] + (i - 1)*1.5, y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,])
    solDataFrame$time = as.factor(currentTime)
    # solData <- rbind(solData, solDataFrame)
    
    # Update the solutions
    solDataFrame <- data.frame(S=sol[2,], x=sol[3,] + (i - 1)*1.5 + 0.5, y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,], g=sol[9,])
    solDataFrame$time = as.factor(currentTime)
    
    solPlot <- solPlot + geom_path(data=solDataFrame, aes(x, -y, color=time), size=1.0)
    
    # Update the solutions
    solDataFrameReflected <- data.frame(S=sol[2,], x=(i - 1)*1.5 + 0.5-sol[3,], y=sol[4,], nX=sol[5,], nY=sol[6,], theta=sol[7,], m=sol[8,], g=sol[9,])
    solDataFrameReflected$time = as.factor(currentTime)

    solPlot <- solPlot + geom_path(data=solDataFrameReflected, aes(x, -y, color=time), size=1.0)
    
  }
  

}

solPlot <- solPlot +  ylab(TeX('$y$')) +   xlab(TeX('$x$')) +  theme(legend.position="none", panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_blank(), text=element_text(size=18)) +
  scale_colour_manual(values=colorRampPalette(brewer.pal(9, "BrBG"))(13)) +
  scale_x_continuous(expand = c(0, 0), limits=c(0, length(sigmaValues)*1.5))

solPlot
```
